Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data

### 2022-01-07 Manuscript/R&R/00-response_letter.Rmd

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\hrule
\vspace*{0.02cm}
\hrule
\vspace*{1cm}
\Large __Response to Reviewers' Comments__
\vspace*{1cm}
\hrule
\vspace*{0.5cm}
\begin{tabular}{llp{11.5cm}}
Manuscript No. & : & SIM-21-0956 \\
Title & : & Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data\\
\end{tabular}
\vspace*{0.5cm}
\hrule
\vspace*{0.02cm}
\hrule

\vspace*{1cm}

First, we would like to thank you for handling this paper and thank the AE and both reviewers for their valuable comments. We have carefully addressed all the comments. Point-by-point responses are given below, where each comment is quoted in _italics_.

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item xxx
    \item yyy
    \item zzz
\end{itemize}

\newpage
# Response to Reviewer \#1
```{r reviewer1, child="01-reviewer_1.Rmd"}
```

\newpage
# Response to Reviewer \#2
```{r reviewer2, child="02-reviewer_2.Rmd"}
```

\newpage
# Response to Reviewer \#4
```{r reviewer4, child="03-reviewer_4.Rmd"}
```

\newpage
# Reference

### 2022-01-09 Manuscript/R&R/02-reviewer_2.Rmd
Update R&R

_The authors claim that the proposed model has the advantage of_
_(1) does not induce excess shrinkage while still estimating a smooth function,_
_(2) allow us to deal with linear/non-linear terms separately and understand if nonlinear terms are necessary_
_(3) more scalable than previous algorithms which require MCMC for model fitting_

_While targeting an important question, the novelty of the solution provided by the authors seem a little bit limited. In addition, the key and most significant step is the re-parametrization, yet there are a few points which I do not completely understand and might need further clarification from the authors:_
_(1) It seems that after re-parametrization, the authors do not further consider the smoothing penalty (Equation (1)) and only include the spike-and-slab LASSO prior. I wonder how does this affect the smoothness of the fitted function and how does it resolves the issue of excess shrinkage._
_(2) I was wondering why the authors choose spike-and-slab LASSO prior in particular. It is claimed in the paper that this prior yields a fast coordinate-descent algorithm, yet coordinate-descent seems also available to other priors as well. I think more discussion on the properties of SSL should be helpful here._
_(3) Since the authors have already separated linear and nonlinear effects, and we might expect linear effects to enter the model first, I wonder why the authors does not incorporate this explicitly into their model by, say, changing the prior for gamma’s to impose different sparsity._

_Besides, some writing of this paper should better be polished and some typos corrected. For example:
(1) In page 3, line 31, what does "in the prior density function" mean?
(2) Same page, line 43 to line 45, the notation here is a little bit confusing.
(3) Page 5, line 17-18, how shall we understand this statement and any explanations?
(4) Page 5 line 37 and page 6 line 21-22, the implication seems to be ‘SSL is not a continuous prior’, which is not true.
(5) Page 6, line 16, missing a period at the end.
(6) Page 6, line 30, there are duplicate "the’".
(7) Page 6, line 56, ‘converge’ should be ‘converges’.
(8) Page 9, line 24-25, there seems to be duplicated "variance-covariance matrix".
(9) page 11, line 8, there are some problems with the table reference.
(10) In section 4, what is definition for the "deviance"?_

>>>>

_The authors claim that the proposed model has the advantage of_

  1. _does not induce excess shrinkage while still estimating a smooth function,_
  2. _allow us to deal with linear/non-linear terms separately and understand if nonlinear terms are necessary_
  3. _more scalable than previous algorithms which require MCMC for model fitting_

_While targeting an important question, the novelty of the solution provided by the authors seem a little bit limited. In addition, the key and most significant step is the re-parametrization, yet there are a few points which I do not completely understand and might need further clarification from the authors:_

_(1) It seems that after re-parametrization, the authors do not further consider the smoothing penalty (Equation (1)) and only include the spike-and-slab LASSO prior. I wonder how does this affect the smoothness of the fitted function and how does it resolves the issue of excess shrinkage._

__Response__:
\bigskip

_(2) I was wondering why the authors choose spike-and-slab LASSO prior in particular. It is claimed in the paper that this prior yields a fast coordinate-descent algorithm, yet coordinate-descent seems also available to other priors as well. I think more discussion on the properties of SSL should be helpful here._

__Response__:
\bigskip

_(3) Since the authors have already separated linear and nonlinear effects, and we might expect linear effects to enter the model first, I wonder why the authors does not incorporate this explicitly into their model by, say, changing the prior for gamma’s to impose different sparsity._

__Response__:
\bigskip

\em
Besides, some writing of this paper should better be polished and some typos corrected. For example:
\begin{enumerate}
\item In page 3, line 31, what does "in the prior density function" mean?
\item Same page, line 43 to line 45, the notation here is a little bit confusing.
\item Page 5, line 17-18, how shall we understand this statement and any explanations?
\item Page 5 line 37 and page 6 line 21-22, the implication seems to be ‘SSL is not a continuous prior’, which is not true.
\item Page 6, line 16, missing a period at the end.
\item Page 6, line 30, there are duplicate "the’".
\item Page 6, line 56, ‘converge’ should be ‘converges’.
\item Page 9, line 24-25, there seems to be duplicated "variance-covariance matrix".
\item page 11, line 8, there are some problems with the table reference.
\item In section 4, what is definition for the "deviance"?
\end{enumerate}
\em

__Response__:
\bigskip

### 2022-01-09 Manuscript/R&R/03-reviewer_4.Rmd
Update R&R

## _2. Main Comments and Questions_
_The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors brie
y allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "\[include] the linear component in the model when non-linear component is selected."_

\em 
Could the authors elaborate on how they do this? I think this point should also be
stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced. One potential downside of
using the double exponential priors on individual basis coefficients, as opposed to the
group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are
regularized. So it is conceivable that BHAM would end up including the nonlinear
component but not the linear one, which is not sensible. How can strong hierarchy
be enforced to ensure that the linear component is always included if the nonlinear
one is selected (but not necessarily the other way around)? This should be detailed in
the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce
strong hierarchy.

For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).

2. Although the authors claim that BHAM is beneficial because it can distinguish linear
effects from nonlinear effects (therefore, we can tell if nonlinear effects are necessary),
this touted benefit is not well-illustrated in the Simulation Study (Section 3) or the
metabolomics data (Section 4). In particular, the simulation setting in Section 3.1
contains only one linear function. The authors should expand their simulation study
to include a greater number of linear functions (in addition to a few nonlinear ones)
to show how well BHAM is truly distinguishing the linear effects from the nonlinear
effects.

It would also be useful to report things like FDR, FNR, Matthews correlation coef-
ficient, etc. for the function selection to see how well the proposed BHAM model
is performing with respect to selection of the nonzero effects, especially compared to
other methods like SBGAM or spikeslabGAM. In addition, if the authors could re-
port a performance metric of how well BHAM is distinguishing nonlinear from linear
functions, e.g. possibly reporting the proportion of simulations where the linear func-
tion (resp. nonlinear) was correctly identified as linear (resp. nonlinear), that would
provide empirical support for BHAM over existing methods.

3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
should also compare it to the method spikeslabGAM of [3]. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM.

4. The authors introduce the EM-IWLS method which can be used for inference of the
regression coefficients. Does this method work for p > n? Or is it only limited to the
case where p $\leq$ n? Even if this algorithm is limited to only the p $\leq$ n case, the benefit
of inference is not illustrated anywhere in the manuscript. Could the authors report
how to use the standard errors from the error variance-covariance matrix returned from
EM-IWLS to construct pointwise confidence intervals? There should be a section in
the manuscript for Uncertainty Quantification which explains how to use the results
from the EM-IWLS algorithm for the regression coefficients to construct pointwise
confidence intervals for the univariate functions themselves.

In addition, the coverage probability of the 95% pointwise intervals constructed by
BHAM in simulations should also be reported. Furthermore, the methods mgcv and
spikeslabGAM also return interval estimates, so the authors would do well to compare
the coverage probability of the confidence intervals constructed by BHAM vs. the
confidence/credible intervals returned by mgcv and spikeslabGAM.

5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
as a DAG. However, there are no other figures in the paper, e.g. plots of function
estimates from BHAM. The authors may consider adding a plot of the function esti-
mates vs. the ground truth in the simulation study (Section 3) for one of the nonlinear
functions and one of the linear functions (these could be, for instance, the left panel
and right panel of a figure). This will illustrate that BHAM can accurately recover
both nonlinear functions and linear functions.

The authors may also give one or two plots from the real data analysis in Section 4 of
some of the significant effects that they discovered. These plots should contain not just
the point estimates but also the the pointwise confidence intervals from the EM-IWLS
method. This will illustrate the benefit of EM-IWLS since it allows practitioners to
assess the uncertainty, in addition to an estimate of the function itself.

## _3. Minor Comments_
1. In Section 1, the authors say that nonparametric methods are an alternative to "black
box" methods like random forest. This is a bit misleading, since the "black box"
methods (random forest, deep learning, etc.) are als nonparametric methods. The
authors might consider rephrasing it (e.g. they could say that semiparametric methods
like GAMs allow for easier interpretation and selection of the significant effects than
the fully nonparametric "black box" methods).

>>>>

## _2. Main Comments and Questions_
\em 1. The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors briefly allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "[include] the linear component in the model when non-linear component is selected."

Could the authors elaborate on how they do this? I think this point should also be stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced. One potential downside of using the double exponential priors on individual basis coefficients, as opposed to the group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are regularized. So it is conceivable that BHAM would end up including the nonlinear component but not the linear one, which is not sensible. How can strong hierarchy be enforced to ensure that the linear component is always included if the nonlinear one is selected (but not necessarily the other way around)? This should be detailed in the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce strong hierarchy.

For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).\em

__Response__:
\bigskip

\em 2. Although the authors claim that BHAM is beneficial because it can distinguish linear
effects from nonlinear effects (therefore, we can tell if nonlinear effects are necessary),
this touted benefit is not well-illustrated in the Simulation Study (Section 3) or the
metabolomics data (Section 4). In particular, the simulation setting in Section 3.1
contains only one linear function. The authors should expand their simulation study
to include a greater number of linear functions (in addition to a few nonlinear ones)
to show how well BHAM is truly distinguishing the linear effects from the nonlinear
effects.

It would also be useful to report things like FDR, FNR, Matthews correlation coef-
ficient, etc. for the function selection to see how well the proposed BHAM model
is performing with respect to selection of the nonzero effects, especially compared to
other methods like SBGAM or spikeslabGAM. In addition, if the authors could re-
port a performance metric of how well BHAM is distinguishing nonlinear from linear
functions, e.g. possibly reporting the proportion of simulations where the linear func-
tion (resp. nonlinear) was correctly identified as linear (resp. nonlinear), that would
provide empirical support for BHAM over existing methods.\em

__Response__:
\bigskip

\em 3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
should also compare it to the method spikeslabGAM of [3]. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM.\em

__Response__:
\bigskip

\em 4. The authors introduce the EM-IWLS method which can be used for inference of the
regression coefficients. Does this method work for p > n? Or is it only limited to the
case where p $\leq$ n? Even if this algorithm is limited to only the p $\leq$ n case, the benefit
of inference is not illustrated anywhere in the manuscript. Could the authors report
how to use the standard errors from the error variance-covariance matrix returned from
EM-IWLS to construct pointwise confidence intervals? There should be a section in
the manuscript for Uncertainty Quantification which explains how to use the results
from the EM-IWLS algorithm for the regression coefficients to construct pointwise
confidence intervals for the univariate functions themselves.

In addition, the coverage probability of the 95% pointwise intervals constructed by
BHAM in simulations should also be reported. Furthermore, the methods mgcv and
spikeslabGAM also return interval estimates, so the authors would do well to compare
the coverage probability of the confidence intervals constructed by BHAM vs. the
confidence/credible intervals returned by mgcv and spikeslabGAM.\em

__Response__:
\bigskip

\em 5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
as a DAG. However, there are no other figures in the paper, e.g. plots of function
estimates from BHAM. The authors may consider adding a plot of the function esti-
mates vs. the ground truth in the simulation study (Section 3) for one of the nonlinear
functions and one of the linear functions (these could be, for instance, the left panel
and right panel of a figure). This will illustrate that BHAM can accurately recover
both nonlinear functions and linear functions.

The authors may also give one or two plots from the real data analysis in Section 4 of
some of the significant effects that they discovered. These plots should contain not just
the point estimates but also the the pointwise confidence intervals from the EM-IWLS
method. This will illustrate the benefit of EM-IWLS since it allows practitioners to
assess the uncertainty, in addition to an estimate of the function itself.\em

__Response__:
\bigskip

## _3. Minor Comments_
\em 1. In Section 1, the authors say that nonparametric methods are an alternative to "black
box" methods like random forest. This is a bit misleading, since the "black box"
methods (random forest, deep learning, etc.) are als nonparametric methods. The
authors might consider rephrasing it (e.g. they could say that semiparametric methods
like GAMs allow for easier interpretation and selection of the significant effects than
the fully nonparametric "black box" methods).\em

__Response__:
\bigskip
